Exploiting Array Syntax in Fortran for Accelerator Programming

### 2011-06-26 papers/GPUScA-2011/performance.tex
Copied IPDPS 2011 paper to GPUScA 2011 workshop (in Galveston)
Near final version.

Performance measurements were made comparing the transformed code with
different versions of the serial shallow-water code.  The serial
versions include two separate Fortran versions:
one using data-parallel notation and the other using explicit looping
constructs.
We also compare with a hand-written OpenCL implementation
that was optimized for local memory usage (no array temporaries).  The
accelerated measurements were made using an NVIDIA Tesla C2050 (Fermi) cGPU with
2.625 GB GDDR5 memory, and 448 processor cores.  The serial
measurements were made using an Intel Xeon X5650 hexacore CPU with 96 GB of RAM
running at 2.67 GHz.  The compilers were gfortran and gcc version 4.4.3
with an optimization level of -O3.

@@

The performance comparison is shown in Figure~\ref{fig:cl-performance} for
varying array sizes (of the state variables).  The time represents an average
of 100 iterations of the outer time-advance loop calling the OpenCL kernel.
This tight loop keeps the OpenCL kernel supplied with threads to take
advantage of potential latency hiding by the NVIDIA GPU.  Any serial code
within this loop would reduce the measured values.

As can be seen in Figure~\ref{fig:cl-performance}, the transformed code achieves very
good results.  It is up to 40 times faster than the best serial Fortran code and
within 25\% of the hand-optimized OpenCL kernel.

\begin{figure}[!t]
\centering
\includegraphics[width=3in]{cl-performance.pdf}
\caption{Performance comparison for varying array size.}
\label{fig:cl-performance}
\end{figure}

>>>>

Performance measurements were made comparing the transformed code with different
versions of the serial shallow-water code.  The serial versions included two
separate Fortran versions: one using data-parallel notation and the other using
explicit looping constructs.  We also compared with a hand-written OpenCL
implementation that was optimized for local memory usage (no array temporaries).
The accelerated measurements were made using an NVIDIA Tesla C2050 (Fermi) cGPU
with 2.625 GB GDDR5 memory, and 448 processor cores.  The serial measurements
were made using an Intel Xeon X5650 hexacore CPU with 96 GB of RAM running at
2.67 GHz.  The compilers were gfortran and gcc version 4.4.3 with an
optimization level of -O3.

@@

%%The performance comparison is shown in Figure~\ref{fig:cl-performance} for varying array sizes (of the state variables).  The time represents an average of 100 iterations of the outer time-advance loop calling the OpenCL kernel. This tight loop keeps the OpenCL kernel supplied with threads to take advantage of potential latency hiding by the NVIDIA GPU.  Any serial code within this loop would reduce the measured values.

Several timing measurements were made by varying the size of the array state
variables.  An average time was obtained by executing 100 iterations of the outer
time-advance loop that called the OpenCL kernel. This tight loop kept the OpenCL
kernel supplied with threads to take advantage of potential latency hiding by
the NVIDIA GPU.  Any serial code within this loop (not present in this study)
would have reduced the measured values.

The transformed code achieved very good results.  In all instances, the
performance of the transformed code was within 25\% of the hand-optimized OpenCL
kernel.  Most of the extra performance of the hand-optimized code can be
attributed to the absence of array temporaries and to packing the three state
variables {\tt H, U}, and {\tt V} into a single vector datatype.

While we did not have an OpenMP code for multi-core comparisons, the transformed
OpenCL code on the NVIDIA C2050 was up to 40 times faster than the best serial
Fortran code executing on the host CPU.

%\begin{figure}[!t]
%\centering
%\includegraphics[width=3in]{cl-performance.pdf}
%\caption{Performance comparison for varying array size.}
%\label{fig:cl-performance}
%\end{figure}
